Improving Estimates of the State of Global Fisheries <br> Depends on Better Data

### 2021-04-24 documents/ovando-etal-assessing-global-fisheries.Rmd
submited version

---
title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
  bookdown::word_document2:
    reference_docx: template.docx
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fish populations play a critical role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain. Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess. But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time. Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

perf <- assess_ram_fits %>% 
  # bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
  mutate(data = as.factor(data)) %>% 
  filter(data != "u_umsy") %>% 
  mutate(data = forcats::fct_recode(data, 
                                    "RAM Index" = "ram-data",
                                    "SAR" = "sar",
                                    "FMI" = "fmi",
                                    "Effective CPUE" = "cpue",
                                    "Effective CPUE+" = "cpue-plus",
                                    "Nominal CPUE" = "nominal-cpue",
                                    "Nominal CPUE+" ="nominal-cpue-plus",
                                    "CMSY" = "cmsy",
                                    "Guess" = "guess",
                                    "RAM F/Fmsy" = "u_umsy"

@@

\newpage
# Introduction {.unnumbered}

@@

What do we currently know about the state of fished marine species around the world? The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals. In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.

@@

Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein). These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being our measure of stock status), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.  @cabral2020 and @gaines2018).

<!-- These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates of *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model. -->

<!-- @costello2016a found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RAM and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning. -->

```{r status-tab, include=FALSE}

@@

state_space_results <- ss_lookup %>% 
  left_join(ss, by = "region") %>% 
  filter(variable == "BvB") %>% 
  group_by(fao_area) %>% 
  filter(year == max(year)) %>% 
  summarise(`B/BMSY RAM` = round(mean(dlm_geomean),2)) %>% 
  ungroup()

@@

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RAM`, everything())

@@

However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the individual unassessed fisheries around the world. The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure. Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries. However, further evaluation of the predictions made by these models [@free2020;@bouch2021;@pons2020] as well as works such as @pauly2013 and @branch2011 call into question this very idea.

In this paper we ask, how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data? We answer this question by aggregating a set of broadly distributed datasets that might be of use to global fishery assessment. We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data. Wecompared these results with estimates from the increasingly utilized (e.g. @palomares2020) default settings of the catch-only version of CMSY described in @froese2017. We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.

@@

The key attribute of sraplus is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner. We paired these data with fisheries in RAM, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of the data types. We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RAM.

@@

At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017. These catch histories can then be augmented with prior information on stock status derived from increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases. The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g. placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores. SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality. While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries. As an alternative, @rousseau2019b present a global reconstruction of fishing effort for countries around the world. We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status. In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).

@@

The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of Winker *et al.* 2018 [@winker2018]. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support. The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status

@@

Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift. When *m* = 2 *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right. The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, Thorson et al. 2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa. We fix the shape parameter based on the distributions reported in Thorson et al. 2012 [@thorson2012] for the genus of the species in question.

$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors. Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model. We deal with this problem by following the methods described in [@winker2018]. to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.

We allow for process error *p_t* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]). Process error *p_t* is assumed to be log-normally distributed, such that

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011]. Unless there is an abundance index to fit to, the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws. For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

@@

where *a_t* is the observed abundance index and *F* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)). When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data. [@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep). Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality. When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population. This is due to the fact that the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the fishing mortality rate is bounded between 0% and 100%). To accommodate this, we generate a catch per effective harvest rate index of abundance, as

@@

In the absence of any data to fit to, the stock reduction algorithm (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See @poole2000 and references therein for a discussion of Borel's Paradox in a fisheries context).

This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood. Borel's Paradox poses a particular problem  for the SRA version of sraplus. Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.

We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin. The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors. See Supplementary Information for a detailed explanation of this problem and our solution.

@@

Simulation testing can be preferable in many ways to comparison against model outputs. However it is not possible to simulation test the value of information derived from empirical relationships between variables such as fisheries management strength and fishery outcomes, which our study depends on. As such we assess model performance through comparison to best available estimates of stock status available in RAM.  We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history. *B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.  Our methods approximated a regional-level assessment exercise, where data beyond catch histories are available at regional levels, but not for specific fisheries.

>>>>

---
title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
  bookdown::word_document2:
    reference_docx: template.docx
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
toc: false
---

@@

perf <- assess_ram_fits %>% 
  # bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
  mutate(data = as.factor(data)) %>% 
  filter(data != "u_umsy") %>% 
  mutate(data = forcats::fct_recode(data, 
                                    "RLSADB Index" = "ram-data",
                                    "SAR" = "sar",
                                    "FMI" = "fmi",
                                    "Effective CPUE" = "cpue",
                                    "Effective CPUE+" = "cpue-plus",
                                    "Nominal CPUE" = "nominal-cpue",
                                    "Nominal CPUE+" ="nominal-cpue-plus",
                                    "CMSY" = "cmsy",
                                    "Guess" = "guess",
                                    "RLSADB F/Fmsy" = "u_umsy"

@@

# Title Page {.unnumbered}

**Title 1**: Improving Estimates of the State of Global Fisheries Depends on Better Data

**Title 2**: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

**Running Title**: Unassessed Fisheries

**Authors**:
Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^

**Affiliations**:

^1^University of Washington, School of Aquatic and Fishery Sciences
1122 NE Boat St, Box 355020 Seattle, WA 98195-5020
Seattle, WA, USA 98103

^2^NOAA Fisheries Alaska Fisheries Science Center, Seattle, WA, USA

^3^Scaleability LLC

^4^Food and Agriculture Organization of the United Nations, Fisheries and Aquaculture,Rome, IT

^5^University of Tasmania, Institute for Marine and Antarctic Studies,Hobart, TAS, AUS

**Authorship**: DO, RH, CM,MR, RS, JT designed model structure and ran analyses. RS, YR, and YE supplied data. All authors contributed to writing of the manuscript

**Conflict of Interest**: RH receives research funding from many groups that have interests in fisheries outcomes including environmental NGOs, foundations, governments and fishing industry groups.

# Abstract {.unnumbered}

Implementation of the United Nations Sustainable Development Goals requires assessments of the global state of fish populations. While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain. Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess. But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time. Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.

# Introduction {.unnumbered}

@@

What do we currently know about the state of fished marine species around the world? The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments summarized in the RAM Legacy Stock Assessment Database (RLSADB) [@ricard2012]. The "assessed" fisheries in RLSADB represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength [@melnychuk2017], trawl footprints [@amoroso2018], and fishing effort [@rousseau2019b], the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals. In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.

@@

Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein). These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being a common measure of stock status), from characteristics of a fishery's catch history.

<!-- These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates of *B/B~MSY~* and other reference points reported in RLSADB are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RLSADB to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model. -->

<!-- @costello2016a found similar rankings of regions in terms of stock status as RLSADB, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RLSADB results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RLSADB and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning. -->

```{r status-tab, include=FALSE, eval=FALSE}

@@

state_space_results <- ss_lookup %>% 
  left_join(ss, by = "region") %>% 
  filter(variable == "BvB") %>% 
  group_by(fao_area) %>% 
  filter(year == max(year)) %>% 
  summarise(`B/BMSY RLSADB` = round(mean(dlm_geomean),2)) %>% 
  ungroup()

@@

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RLSADB`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RLSADB`, everything())

@@

However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating the state of the individual unassessed fisheries around the world. The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure. Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries. However, further evaluation of the predictions made by these models [@free2020;@bouch2021;@pons2020] as well as works such as @pauly2013 and @branch2011 call into question this very idea.

In this paper we ask, how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data? We answer this question by aggregating a set of broadly distributed datasets that might be of use to global fishery assessment. We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data. We compared these results with estimates from the increasingly utilized (e.g. @palomares2020) default settings of the catch-only version of CMSY described in @froese2017. We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.

@@

The basis of this analysis is a new stock assessment software package we call [sraplus](https://github.com/DanOvando/sraplus) (https://github.com/DanOvando/sraplus). sraplus is an extension of stochastic stock reduction analysis (SRA) [@kimura1984;@walters2006], which allows users to combine a biomass dynamics model with a variety of data sources (e.g. priors on recent stock status, or an index of abundance) in order to produce estimates of the state of a fishery over time. The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner. We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of data types. We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RLSADB.

@@

At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017. These catch histories can then be augmented with prior information on stock status, derived from expert opinion, or using built-in prior generating models based on increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases. The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g. placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores. SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality. While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries. As an alternative, @rousseau2019b present a global reconstruction of fishing effort for countries around the world. We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status. In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).

@@

The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of Winker *et al.* 2018 [@winker2018]. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support.

<!-- The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status -->

@@

Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift. When *m* = 2 *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right. The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimates of the ratio of *B~MSY~* to *K* for many fish taxa. We fix the shape parameter based on the distributions reported in @thorson2012 for the genus of the species in question.

$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors. Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model. We deal with this problem by following the methods described in [@winker2018] to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.

We allow for process error $p_t$ (in the manner of the stochastic stock reduction analysis (SRA) suggested by @walters2006). Process error $p_t$ is assumed to be log-normally distributed, such that

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011]. Unless there is an abundance index to fit to, the model runs as an SRA. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws. For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

@@

where $a_t$ is the observed abundance index and *F* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)). When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data. [@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep). Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality. When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population. This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized fishing mortality rate must be between between 0% and 100%). To accommodate this, we generate a catch per effective harvest rate index of abundance, as

@@

In the absence of any data to fit to, sraplus works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history, life history priors, and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See @poole2000 and references therein for a discussion of Borel's Paradox in a fisheries context).

This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood. Borel's Paradox poses a particular problem when there are no data to fit to though (i.e. when the model is simply filtering through prior distributions in the manner of a traditional SRA). Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse. In this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.

We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step sampling-importance-resampling (SIR) algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean probability density *p* (defined by the prior distributions of estimated parameters) of each bin.

$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$

We then divide the mean probability density of bin *i* evenly among each of the draws within that bin *n*

$$p(n_i) = \frac{p(bin_i)}{N_i}$$

And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.

The net result of this is that it allows users to place an explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior. While the range of possible life history values supplied still influences stock status under this approach, this prior predictive tuning process makes the resulting priors more consistent with explicit priors on recent stock status supplied by the user. Users can turn this functionality off and instead base priors on stock status primarily on life history. See Supplementary Information for a detailed explanation of this problem and our solution.

@@

Simulation testing can be preferable in many ways to comparison against model outputs. However it is not possible to simulation test the value of information derived from empirical relationships between variables such as fisheries management strength and fishery outcomes, which our study depends on. As such we assess model performance through comparison to best available estimates of stock status available in RLSADB.  We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RLSADB, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history. *B/B~MSY~* values from RLSADB are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RLSADB stocks with regional-level SAR, FMI, and effort data.  Our methods approximated a regional-level assessment exercise, where data beyond catch histories are available at regional levels, but not for specific fisheries. We also estimated *B/B~MSY~* values of our candidate RLSADB stocks by using sraplus to fit to an abundance index drawn directly from RLSADB itself "RLSADB Index", as a measure of the ability of models like sraplus if given perfect information
